Finiteness in Cubical Type Theory

### 2020-06-26 sections/search.tex
finished permutations section

What we actually need is not just a function between positions, but an
\emph{isomorphism} between them.
In types:
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{iso-perm}
\end{agdalisting}
Where an isomorphism is defined as follows:
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{isomorphism}
\end{agdalisting}
\todo{Should this isomorphism definition be put earlier in the intro with the
  equivalences etc?}
While it may look complex, this term is actually composed of individual
components we've already proven finite.
First we have \(\AgdaDatatype{Fin}\;n\rightarrow\AgdaDatatype{Fin}\;n\):
functions between finite types are, as we know, finite
(Theorem~\ref{split-enum-pi}).
We take a pair of them: pairs of finite things are \emph{also} finite
(Lemma~\ref{split-enum-sigma}).
To get the next two components we can filter to the subobject: this requires
these predicates to be decidable.
We will construct a term of the following type:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{dec-type}
\end{agdalisting}
So can we construct such a term? Yes!

@@

\subsection{Parenthesising}
\subsection{Putting It All Together}
\section{Automating Proofs}
One use for above constructions is the automation of certain proofs.
In \cite{firsovDependentlyTypedProgramming2015}, which uses a similar approach
to ours, the \(\AgdaDatatype{Pauli}\) group is used as an example.
\ExecuteMetaData[agda/Data/Pauli.tex]{def}
As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
proofs prohibitively verbose.

>>>>

What we actually need is not just a function between positions, but an
\emph{isomorphism} between them.
In types:
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{iso-perm}
\end{agdalisting}
Where an isomorphism is defined as follows:
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{isomorphism}
\end{agdalisting}
\todo{Should this isomorphism definition be put earlier in the intro with the
  equivalences etc?}
While it may look complex, this term is actually composed of individual
components we've already proven finite.
First we have \(\AgdaDatatype{Fin}\;n\rightarrow\AgdaDatatype{Fin}\;n\):
functions between finite types are, as we know, finite
(Theorem~\ref{split-enum-pi}).
We take a pair of them: pairs of finite things are \emph{also} finite
(Lemma~\ref{split-enum-sigma}).
To get the next two components we can filter to the subobject: this requires
these predicates to be decidable. \todo{Need to do filter subobject in topos section}
We will construct a term of the following type:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/ManifestBishop.tex]{dec-type}
\end{agdalisting}
So can we construct such a term? Yes!

@@

Unfortunately this implementation is too slow to be useful.
As nice and declarative as it is, fundamentally it builds a list of all possible
pairs of functions between \(\AgdaDatatype{Fin}\;n\) and itself (an operation
which takes in the neighbourhood of \(\mathcal{O}(n^n)\) time), and then tests
each for equality (which is likely worse than \(\mathcal{O}(n^2)\) time).
We will instead use a factoriadic encoding: this is a relatively simple encoding
of permutations which will reduce our time to a blazing fast
\(\mathcal{O}(n!)\).
It is expressed in Agda as follows:
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{perm-def}
\end{agdalisting}
It is a vector of positions, each represented with a \(\AgdaDatatype{Fin}\).
Each position can only refer to the length of the tail of the list at that
point: this prevents two input positions mapping to the same output point, which
was the problem with the naive encoding we had previously.
And it also has a relatively simple proof of finiteness:
\begin{agdalisting}
  \ExecuteMetaData[agda/Countdown.tex]{perm-fin}
\end{agdalisting}
\subsection{Parenthesising}
Our next step is figuring out a way to encode the parenthesisation
\subsection{Filtering}
\subsection{Putting It All Together}
\section{Automating Proofs}
One use for above constructions is the automation of certain proofs.
In \cite{firsovDependentlyTypedProgramming2015}, which uses a similar approach
to ours, the \(\AgdaDatatype{Pauli}\) group is used as an example.
\ExecuteMetaData[agda/Data/Pauli.tex]{def}
As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
proofs prohibitively verbose.
